An Event-based Architecture for Multi-population Optimization Algorithms

### 2020-01-21 draft.tex
New section

The organization of the paper is as follows. First, in Section \ref{multi}, we present a
background of the fundamental issues of integrating nature-inspired optimization
and multi-population methods. Section \ref{soa} presents state of the art relevant to
our work. In Section \ref{method}, we present the proposed method.  Section \ref{setup} describes
the design of the empirical evaluation we designed to assess the effectiveness
of the method, and in Section \ref{results}, we report and discuss the results. Finally, we
offer the conclusions of this paper and suggestions on future work in Section \ref{conclusions}.

@@

\subsection{Comparison} 
\label{comparison}

Next, we compare the message-driven architecture we presented against four other
proposals for multi-population algorithms found in the literature. We center the
comparison on the coupling and communication between the main components:
Subpopulations, Processing Nodes, and Algorithms.  In Figure \ref{fig:population_message} we show the
main components of a message-driven architecture. We can see that the algorithm
and the subpopulations are separated, and the Population Queue keeps the state
of subpopulations. While algorithms only need the subpopulations as parameters
for their execution.  In contrast, in the classic Island model shown in Figure
\ref{fig:classicisland}, algorithms are coupled with a subpopulation, and the same processing
node keeps both the algorithm process and the state of the subpopulation.  Other
processing nodes follow the same configuration and form of execution, and they
only pass specific individuals between them.  In this model, adding additional
processing nodes, on-runtime, can be more difficult, because they need to have
an addressing mechanism. % JJ, help I don't know if this is clearly understood, a mean like a DNS - Mario

@@

@@

\subsection{Workers} 
\label{edm}

>>>>

The organization of the paper is as follows. First, in Section \ref{multi}, we present a
background of the fundamental issues of integrating nature-inspired optimization
and multi-population methods. Section \ref{soa} presents state of the art relevant to
our work. In Section \ref{method}, we present the proposed method and the container-based
application in Section \ref{docker} .  Section \ref{setup} describes
the design of the empirical evaluation we designed to assess the effectiveness
of the method, and in Section \ref{results}, we report and discuss the results. Finally, we
offer the conclusions of this paper and suggestions on future work in Section \ref{conclusions}.

@@

\subsection{Comparison with other cloud-based works} 
\label{comparison}

Next, we compare the message-driven architecture we presented against four other
proposals for multi-population algorithms found in the literature. We center the
comparison on the coupling and communication between the main components:
Subpopulations, Processing Nodes, and Algorithms.  In Figure \ref{fig:population_message} we show the
main components of a message-driven architecture. We can see that the algorithm
and the subpopulations are separated, and the Population Queue keeps the state
of subpopulations. While algorithms only need the subpopulations as parameters
for their execution.

% We need to add references to the respective papers - Mario

In contrast, in the classic Island model shown in Figure
\ref{fig:classicisland}, algorithms are coupled with a subpopulation, and the same processing
node keeps both the algorithm process and the state of the subpopulation.  Other
processing nodes follow the same configuration and form of execution, and they
only pass specific individuals between them.  In this model, adding additional
processing nodes, on-runtime, can be more difficult, because they need to have
an addressing mechanism. % JJ, help I don't know if this is clearly understood, a mean like a DNS - Mario

@@

A typical design pattern used to alleviate the above drawbacks is the use of a
central repository or pool of individuals that is available to all processing
nodes.  In Figure \ref{fig:pool_island} we show the main components of a pool-based
multi-population algorithm. Although algorithms and subpopulation state remains
coupled, now, processing nodes do not communicate directly with each other.
Instead, they interchange individuals with the pool. Communication between nodes
is not affected if the system adds or removes nodes because they do not know
about each other.  Since the state of the global population is stored in
external processing nodes, the system needs additional communication and
processing for keeping track of each subpopulation.

We show another pool-based approach in Figure  \ref{fig:population_pool}. In
this design, the global population is stored in a central repository, while
isolated algorithms take random samples of the global population and use this
temporal subpopulation as parameters. An advantage of this design is that the
sampling provides a type of migration between isolated subpopulations. A problem
found is that when a processing node returns a subpopulation to the pool, the
state of the subpopulation is lost. But having the algorithm decoupled opens the
possibility of having a serverless-function implementation.

@@

\section{Container-based application} 
\label{docker}

We have mentioned several advantages of a cloud-native application for running
multi-population based algorithms, but there are in our experience several
practical issues of this approach. In general, cloud platforms have free-tiers
for their services, and many offer academic discounts to students. But once we
pass to the pay-as-you-go tier, computing resources need to be managed.  First,
in many cases, there is a credit-card need to guarantee the payment of the
resources consumed and someone responsible for assigning quotas and, more
importantly, special keys for using the cloud's API.  If these keys are
compromised or shared by their holders,  additional costs or administrative
procedures are in order. For those cases when a prototype platform is needed,
for instance, students developing a new algorithm variant or experimenting with
a published work, a container-based platform can be more practical.

\subsection{Containers} 
\label{edm}

\subsection{Docker compose} 
\label{edm}
